\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-25 k-planar.tex
Some definitions around embedded graphs

We first formally define $k$-planar graphs.  An \defin{embedded graph} $G$ is a graph with $V(G)\subset\R^2$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$. The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.  When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.  The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\R^2$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior. The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$. A \defin{crossing} in an embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common. Without any loss of generality we may assume that the interiors of any two edges of $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.  This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.  If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$ then we say that $vw$ and $xy$ \defin{cross at} $p$.

@@

The definition of $k$-planar graphs naturally generalises to genus-$g$ surfaces. A graph $G$ embedded on a surface $\Sigma$ is \defin{$(\Sigma,k)$-plane} if every edge of $G$ is involved in at most $k$ crossings.  A graph $G$ is \defin{$(g,k)$-planar} if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$. \cref{AddDummy} immediately generalises as follows:

@@

\begin{proof}[Proof of \cref{d_framed_product_stucture}]
	Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$
	\note{DW}{Replace ``outer face'' by ``outer-cycle'' since a face is a defined to be a portion of the plane. }.	Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.  Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.  Now apply \cref{induction} on $G$, $G_0$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.  Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, $H$ is planar.

>>>>

We first formally define $k$-planar graphs.  For a surface $\Sigma$,  an \defin{$\Sigma$-embedded graph} $G$ is a graph with $V(G)\subset\Sigma$ in which each edge $vw\in E(G)$ is a curve\footnote{A \defin{curve} in a surface $\Sigma$ is a continuous function $f:[0,1]\to \Sigma$. The points $f(0)$ and $f(1)$ are called the \defin{endpoints} of the curve.  When there is no danger of misunderstanding we treat a curve $f$ as the point set $\{f(t):0\le t\le 1\}$.  The \defin{interior} of $f$ is the point set $\{f(t):0<t<1\}$.} in $\Sigma$ with endpoints $v$ and $w$ and not containing any vertex of $G$ in its interior. The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$. A \defin{crossing} in a $\Sigma$-embedded (multi)graph $G$ is an unordered pair of distinct edges $\{vw,xy\}\subseteq E(G)$ whose interiors have a point in common.

If an embedded graph $G$  has no crossings then it is \defin{non-crossing} graph.  When $G$ is non-crossing and $F$ is a face of $G$ then $V(F)$ denotes the set of vertices in $G$ on the boundary of $F$.  When $F$ is homeomorphic to a disk, the \defin{facial walk} of $F$ is the cycle consisting of the edges and vertices of $G$ on the boundary of $F$.  When $\Sigma=\R^2$ is the Euclidean plane and each edge of $G$ is involved in at most $k$ crossings then $G$ is a $k$-plane graph.  A graph $G$ is \defin{$k$-planar} if is isomorphic to some $k$-plane graph.  A $0$-plane graph is a \defin{plane} graph and a $0$-planar graph is a \defin{planar} graph.

Without any loss of generality we may assume that the interiors of any two edges of a $\sigma$-embedded graph $G$ have at most one point in common and that no point in $\R^2$ is contained in the interior of three or more edges of $G$.  This assumption can be enforced by local changes that do not change the number of crossings each edge is involved in.  If $p$ is a common point in the interior of two crossing edges $vw$ and $xy$ then we say that $vw$ and $xy$ \defin{cross at} $p$.

@@

The definition of $k$-planar graphs naturally generalises to genus-$g$ surfaces. A graph $\Sigma$-embedded graph $G$ is \defin{$(\Sigma,k)$-plane} if every edge of $G$ is involved in at most $k$ crossings.  A graph $G$ is \defin{$(g,k)$-planar} if it is isomorphic to some $(\Sigma,k)$-plane graph, for some surface $\Sigma$ with Euler genus at most $g$. \cref{AddDummy} immediately generalises as follows:

@@

\begin{proof}[Proof of \cref{d_framed_product_stucture}]
	Let $G$ be a $d$-framed multigraph with frame $G_0$ having outer face $F_0$
	% \note{DW}{Replace ``outer face'' by ``outer-cycle'' since a face is a defined to be a portion of the plane. }.
  %\note{PM}{I added some conventions around the definition of $\Sigma$-embedded graph so that $V(F)$ is defined for a face $F$ in a non-crossing embedded graph.}
  Let $T$ be a BFS forest of $G_0$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.  Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.  Now apply \cref{induction} on $G$, $G_0$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.  Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, $H$ is planar.
